\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-04-24 doc/incertae_sedis.tex
Reformating to be readable with preview-latex.

We seek a semantics for supertrees with \emph{incertae sedis }taxa
that satisfy the following properties:
\begin{enumerate}
    \item An \emph{incertae sedis} node may intrude into its siblings and their
descendants.
    \item Two \emph{incertae sedis} siblings may not be
interdigitated.
    \item Any placement $T\to T^{*}$ of a tree $T$ with
\emph{incertae sedis} taxa is also accessible from a tree $T^{\prime}$
that is created from $T$ by taking a taxon, attaching it to an
ancestor, and marking it \emph{incertae sedis}. \mthcomment{MTH: I don't understand this.}
\bdrcomment{BDR: Figure \ref{fig:Placing-one-incertae} is an illustration of this.  Suppose you have a genus that is incertae sedis and a family that is incertae sedis and attached at the kingdom level.  We could in theory place the family (.e.g $I$ in the figure) at the right level.  It then sits next to the incertae sedis genus (e.g. $AI$ in the figure).  We could then place the genus within the family.

The goal above is trying to say (and failing to be clear at saying) that if we have a family that is sibling to an incertae sedis genus, then the placement of the genus within the family is still a possibility if we reattach the family as a child of a kingdom node and mark it incertae sedis.
}
    \item The semantics of an
\emph{incertae sedis} node does not depend on assigned ranks, but only
on the tree.
    \item Semantics is based on deriving a split for each
branch of the tree.
\end{enumerate}
Here we focus on a split-based
semantics instead of a semantics based on apply of a series of edit
operations to the tree.
The split-based semantics allows us to seek
trees that are simultaneously consistent with a set of split-
constraints.
The use of a split-based semantics does impose some
limitations.
These come primarily because in the edit-based semantics,
after we place an \emph{incertae sedis} taxon, it is no loner
\emph{incertae sedis}, whereas with the split-based semantics taxa are
either always \emph{incertae sedis} or never \emph{incertae sedis}.

This has an effect, for example, when we consider two incertae sedis
siblings $A$ and $B$.
If we can place $A$ within $B$, and $B$ within
$A$, then a split- based semantics allows us to freely inter-digitate
$A$ and $B$.
If we do not want to allow $A$ and $B$ to be inter-
digitated, then we cannot both $A$ within $B$ and $B$ within $A$.
In
addition, we can imagine that a family $A$ is incertae sedis within a
kingdom.
In an edit-based semantics, we could place $A$ as a sibling
to a genus $B$, mark $A$ as non-\emph{incertae-sedis}, and then place
$B$ within $A$.
In a split-based semantics, we must disallow either
$B$ within $A$ or $A$ within $B$ if we want to disallow
interdigitation.

As a result of this, we cannot satisfy all five of these criteria.
Thus, we describe two alternative split-based semantics.
The non-
interdigitation semantics gets \# (1,2,4,5), and the interdigitation
semantics gets (1,3,4,5).
This makes the point that each semantics
results from different choices about which criteria to satisfy.

@@

\emph{}

>>>>

Here we focus on a split-based semantics instead of a semantics based on apply
of a series of edit operations to the tree. The split-based semantics allows us
to seek trees that are simultaneously consistent with a set of split-constraints.
The use of a split-based semantics does impose some limitations. These come
primarily because in the edit-based semantics, after we place an
\emph{incertae sedis} taxon, it is no loner \emph{incertae sedis}, whereas with
the split-based semantics taxa are either always \emph{incertae sedis} or never
\emph{incertae sedis}.

This has an effect, for example, when we consider two incertae sedis siblings $A$
and $B$. If we can place $A$ within $B$, and $B$ within $A$, then a split- based
semantics allows us to freely inter-digitate $A$ and $B$. If we do not want to
allow $A$ and $B$ to be inter-digitated, then we cannot both $A$ within $B$ and $B$
within $A$. In addition, we can imagine that a family $A$ is incertae sedis within a
kingdom. In an edit-based semantics, we could place $A$ as a sibling to a genus $B$,
mark $A$ as non-\emph{incertae-sedis}, and then place $B$ within $A$. In a split-based
semantics, we must disallow either $B$ within $A$ or $A$ within $B$ if we want to disallow
interdigitation.

As a result of this, we cannot satisfy all five of these criteria. Thus, we
describe two alternative split-based semantics. The non-interdigitation
semantics gets \# (1,2,4,5), and the interdigitation semantics gets (1,3,4,5).
This makes the point that each semantics results from different choices about
which criteria to satisfy.

@@
